Exploratory Analysis

### 2023-05-01 code/analysis_code/exploratory_analysis.qmd

# Setup
The `exploratoryanalysis.r` script needs to be indicated as the file from which to pull code chunks from. 
```{r, include=FALSE, cache=FALSE}
knitr::read_chunk('exploratoryanalysis.R')
```

Next, load the needed packages.
```{r, packages, message = FALSE, warning = FALSE}
```

Then, load the data from the `processed_data` folder.
```{r, loaddata}
```

@@

I would like to look at each variable individually to see an overview of the indicator. These summary tables will be saved in the `results` folder. Similar to the cleaning code, I will separate the summary tables into groups of indicators. 
```{r, exploratorytables1}
```
The BCG coverage indicator has 4 dimensions and 12 subgroups and appears to have mostly data from smaller population sizes (most of the data has less than 720) and high BCG coverage (mean of 88.7%).

@@

TB mortality is similar to the incidence variable in terms of subgroups, year, and population. The mortality outcome is also continuous, and the highest mortality rate is 140 deaths/100000 population.
```{r, exploratorytables3}
```
Prevalence to notification ratio studies 2 subgroups within a dimension. There does not appear to be a trend toward shorter or longer length of time.

@@

Attitudes toward keeping TB infection a secret studies 2 subgroups within a dimension for the overall measure, and when split into female and male, measures 4 dimensions with 14 subgroups. All data was collected from 2011 to 2019, and each group tends to have a relatively low percentage (means of 23.3%, 26.8%, and 19.9%).
```{r, exploratorytables5}
```
Knowledge indicators of people reporting TB is spread through cough has similar dimensions and subgroups as the attitudes indicators. Data collection also occured during the same years (2011-2019). The groups all tended to have similar levels of knowledge based on the overall view (means of 66.7%, 62.3%, 68.8%).

@@

```{r, exploratoryfigures1}
```
The first plot show BCG coverage among under 1 year olds by economic status, education, place of residence, and sex. The economic status and education groups shows a clear trend of high coverage for higher income and higher education. The place of residence shows a higher coverage for urban areas, but the rural sites show greater spread of data. The coverage seems about equal between females and males. all categories show a significant number of outliers.

The second plot shows catastrophic costs due to TB by economic level and TB resistance. Higher income seems to be correlated with lower catastrophic costs, and drug resistance has a much higher correlation with costs compared to non-drug resistant TB. 
```{r, exploratoryfigures2}
```
The third plot shows case detection rate by age and sex. Higher age groups tend to have higher detection rates (although the age group is quite large).

The forth plot shows drug resistance by sex. Neither group appears to have a higher percentage of drug resistance.
```{r, exploratoryfigures3}
```
The 5a plot and 5b plots show TB incidence by country, separated by sex for the high burden countries and full data set. The countries are ordered from lowest to highest rate, and males seem to have a higher proportion of TB cases.The `highburden` data set appears to include most of the countries that have at least one new case per 100,000.

@@

The 8a and 8b plots show TB prevalence by country, showing differences between rural and urban areas for the `highburden` and full data set. There appears to be a small trend of urban areas having higher prevalence then rural. 
```{r, exploratoryfigures4}
```
The ninth plot shows the overall attitude towards keeping a family member's TB a secret by sex. Females tend to be correlated with a higher percentage compared to males.

@@

```{r, exploratoryfigures5}
```
The twelfth plot shows TB knowledge about spreading the disease through cough by sex. Males tended to have a slightly higher proportion compared to females.

>>>>

# Setup

The `exploratoryanalysis.r` script needs to be indicated as the file from which to pull code chunks from.

```{r, include=FALSE, cache=FALSE}
knitr::read_chunk('exploratoryanalysis.R')
```

Next, load the needed packages.

```{r, packages, message = FALSE, warning = FALSE}
```

Then, load the data from the `processed_data` folder.

```{r, loaddata}
```

@@

I would like to look at each variable individually to see an overview of the indicator. These summary tables will be saved in the `results` folder. Similar to the cleaning code, I will separate the summary tables into groups of indicators.

```{r, exploratorytables1}
```

The BCG coverage indicator has 4 dimensions and 12 subgroups and appears to have mostly data from smaller population sizes (most of the data has less than 720) and high BCG coverage (mean of 88.7%).

@@

TB mortality is similar to the incidence variable in terms of subgroups, year, and population. The mortality outcome is also continuous, and the highest mortality rate is 140 deaths/100000 population.

```{r, exploratorytables3}
```

Prevalence to notification ratio studies 2 subgroups within a dimension. There does not appear to be a trend toward shorter or longer length of time.

@@

Attitudes toward keeping TB infection a secret studies 2 subgroups within a dimension for the overall measure, and when split into female and male, measures 4 dimensions with 14 subgroups. All data was collected from 2011 to 2019, and each group tends to have a relatively low percentage (means of 23.3%, 26.8%, and 19.9%).

```{r, exploratorytables5}
```

Knowledge indicators of people reporting TB is spread through cough has similar dimensions and subgroups as the attitudes indicators. Data collection also occured during the same years (2011-2019). The groups all tended to have similar levels of knowledge based on the overall view (means of 66.7%, 62.3%, 68.8%).

@@

```{r, exploratoryfigures1}
```

The first plot show BCG coverage among under 1 year olds by economic status, education, place of residence, and sex. The economic status and education groups shows a clear trend of high coverage for higher income and higher education. The place of residence shows a higher coverage for urban areas, but the rural sites show greater spread of data. The coverage seems about equal between females and males. all categories show a significant number of outliers.

The second plot shows catastrophic costs due to TB by economic level and TB resistance. Higher income seems to be correlated with lower catastrophic costs, and drug resistance has a much higher correlation with costs compared to non-drug resistant TB.

```{r, exploratoryfigures2}
```

The third plot shows case detection rate by age and sex. Higher age groups tend to have higher detection rates (although the age group is quite large).

The forth plot shows drug resistance by sex. Neither group appears to have a higher percentage of drug resistance.

```{r, exploratoryfigures3}
```

The 5a plot and 5b plots show TB incidence by country, separated by sex for the high burden countries and full data set. The countries are ordered from lowest to highest rate, and males seem to have a higher proportion of TB cases.The `highburden` data set appears to include most of the countries that have at least one new case per 100,000.

@@

The 8a and 8b plots show TB prevalence by country, showing differences between rural and urban areas for the `highburden` and full data set. There appears to be a small trend of urban areas having higher prevalence then rural.

```{r, exploratoryfigures4}
```

The ninth plot shows the overall attitude towards keeping a family member's TB a secret by sex. Females tend to be correlated with a higher percentage compared to males.

@@

```{r, exploratoryfigures5}
```

The twelfth plot shows TB knowledge about spreading the disease through cough by sex. Males tended to have a slightly higher proportion compared to females.
